---
title: "POLI7002:Homework 4"
author: "Hu, Yue"
date: '`r Sys.Date()`'
output: html_notebook
---

# Data manipulation
Package preload:
```{r}
library(pacman)
p_load(ggplot2, # Visualization
       rjson, quanteda, tm, # analysis
       haven, descr, stringr, data.table, feather, 
       tibble, purrr, dplyr) # data wrangling
```

The `quanteda` is good at Luke dictionary (`liwc`).

Input the dictionary and text files.
In the dictionary, there are multiple categories.
What we are interested in in this project is "positive" and "negative."

```{r}
liwc <- dictionary(file = "LIWC2007.txt",format = "LIWC")
names(liwc)
#load trump download
trump <- fromJSON(file = "../hw3/trump.json")
```

The input `json` is a list of texts and their meta data. 
The following step is thus to draw the texts out and convert it to a `VCorpus`, so as to inputing into `quanteda` functions.

```{r}
trump_articles <- map(trump, "text")
trump_corpus <- Corpus(VectorSource(trump_articles)) %>% corpus()

# count the words
trump_wc <- dfm(trump_corpus) %>% # creating a document-feature matrix
  rowSums()
head(trump_wc)
```

Then we can use the dictionary file to analyze the `document-feature matrix` (viz., the DTM) just got.

```{r}
trump_dfm <- dfm(trump_corpus,  # corpus
                 dictionary = liwc, # dictionary file
                 removeNumbers = TRUE,  # remove the digits
                 removePunct = TRUE,  # remove the punctuations
                 ignoredFeatures = stopwords("english")) # tokenization
```

What we are looking for are the "positive" and "negative" categories.
So, we search where the root *pos* and *neg* locate according to the index stored in the `trump_dfm@Dimnames$features`.
Here we use the basic `grep` function, since it gives the position directly.
`stringi::str_locale_get` seems having the same function---but, why bothers~

```{r}
grep("pos", trump_dfm@Dimnames$features, value = T)
grep("neg", trump_dfm@Dimnames$features, value = T)
```

So, the goal here is to create a bar chart in R that shows the percentage of positive and negative words for each month in your data.

```{r eval = F}
#liwc
trump_dfm <- dfm(trump_corpus, 
                 dictionary = liwc,
                 removeNumbers = TRUE, 
                 removePunct = TRUE, 
                 ignoredFeatures = stopwords("english"))
trump_pos<-as.numeric(trump_dfm[,24])
trump_neg<-as.numeric(trump_dfm[,25])

#boxplot
boxplot(data.frame(cbind("Positive"=trump_pos/trump_wc,"Negative"=trump_neg/trump_wc)))

#custom dictionary
my_dictionary <- list(spending=c("expensive", "cost", "deficit","spending"),
               spending2=c("expensive", "cost", "deficit","spend*"))

trump_dfm2<-dfm(trump_corpus,dictionary=my_dictionary,removeNumbers = TRUE, removePunct = TRUE, ignoredFeatures = stopwords("english"))
trump_spend<-as.numeric(trump_dfm2[,1])
trump_spend2<-as.numeric(trump_dfm2[,2])

#boxplot...spending
boxplot(data.frame(cbind("Spending1"=trump_spend/trump_wc,"Spending2"=trump_spend2/trump_wc)))

```

